When two-point correlation function fails to describe a structure...

### 2024-03-06 paper.tex
Be more explicit about the Riemann hypothesis

Breuillard et~al. \cite{breuillard2019irreducibility} give the following
approximation for this probability based on extended Riemann hypothesis:
\begin{equation}
  P(\text{$f$ is irreducible}) = 1 - \sqrt{\frac{2}{\pi d}} + O(d^{-1})
  \label{eq:prob-irr}
\end{equation}

@@

Now we get our main result:
\begin{theorem}
  Let $A$ be a 2-dimensional $D_1 \times D_2$ array of 0's and 1's where each
  element is chosen randomly and independently with probability $1/2$ with
  exception of two elements $A[0, 0]$ and $A[D_1 - 1, 0]$, which are 1's. Then
  $A$ does not have non-trivial collisions in the two-point correlation function
  with probability at least
  \begin{equation}
    p = 1 - \sqrt{\frac{2}{\pi (D_1 - 1)}}
    \label{eq:main-result}
  \end{equation}
\end{theorem}
\begin{proof}
  Let $f(x, y)$ be an image of $A$ under Z-transform and a change of variables
  $x = z_1^{-1}, y = z_2^{-1}$ with $\deg_x f = d$, where $d = D_1 - 1$. Let
  $\phi f = f \mod y$ be a ``partial evaluation'' homomorphism from
  $\mathbb{Z}[x, y]$ to $\mathbb{Z}[x]$. Let $f_x(x) = \phi f = f(x,0)$. Under
  conditions of the theorem $f(0) = 1$ and $\deg f_x = d$.  According to
  \cref{eq:prob-irr} we have
  \begin{equation*}
    p\{\text{$f_x$ is irreducible}\} = 1 - \sqrt{\frac{2}{\pi d}} + O(d^{-1})
  \end{equation*}
  Now we need to prove a similar result for $f$.

>>>>

Breuillard et~al. \cite{breuillard2019irreducibility} give the following
approximation for this probability if the Riemann hypothesis holds for the
Dedekind zeta function $\zeta_K$ for all number fields of the form
$K = \mathbb{Q}(a)$, where $a$ is a root of $f$.
\begin{equation}
  P(\text{$f$ is irreducible}) = 1 - \sqrt{\frac{2}{\pi d}} + O(d^{-1})
  \label{eq:prob-irr}
\end{equation}

@@

Now we get our main result:
\begin{theorem}
  Let $A$ be a 2-dimensional $D_1 \times D_2$ array of 0's and 1's where each
  element is chosen randomly and independently with probability $1/2$ with
  exception of two elements $A[0, 0]$ and $A[D_1 - 1, 0]$, which are
  1's. Suppose the Riemann hypothesis for the Dedekind zeta function $\zeta_K$
  holds for all fields $K = \mathbb{Q}(a)$ where $a$ is any root of $f_x(x)$ and
  $f_x(x)$ is an image of the slice $A[0:D_1 -1, 0]$ under Z-transform and a
  change in variables $x = z^{-1}$. Then $A$ does not have non-trivial
  collisions in the two-point correlation function with probability at least
  \begin{equation}
    p = 1 - \sqrt{\frac{2}{\pi (D_1 - 1)}}
    \label{eq:main-result}
  \end{equation}
\end{theorem}
\begin{proof}
  Let $f(x, y)$ be an image of $A$ under Z-transform and a change of variables
  $x = z_1^{-1}, y = z_2^{-1}$ with $\deg_x f = d$, where $d = D_1 - 1$. Let
  $\phi f = f \mod y$ be a ``partial evaluation'' homomorphism from
  $\mathbb{Z}[x, y]$ to $\mathbb{Z}[x]$. Oviously,
  $f_x(x) = \phi f = f(x,0)$. Under conditions of the theorem $f(0) = 1$ and
  $\deg f_x = d$.  According to \cref{eq:prob-irr} we have
  \begin{equation*}
    p\{\text{$f_x$ is irreducible}\} = 1 - \sqrt{\frac{2}{\pi d}} + O(d^{-1})
  \end{equation*}
  Now we need to prove a similar result for $f$.
